\ZpL: a p-adic precision package

### 2018-01-29 ZpL/ZpL.tex
More on absolute precision of an element

\newtheorem{theo}{Theorem}[section]
\newtheorem{lem}[theo]{Lemma}
\newtheorem{prop}[theo]{Proposition}
\newtheorem{cor}[theo]{Corollary}
\newtheorem{quest}[theo]{Question}
\newtheorem{conj}[theo]{Conjecture}
\theoremstyle{definition}
\newtheorem{rmq}[theo]{Remark}
\newtheorem{ex}[theo]{Example}
\newtheorem{deftn}[theo]{Definition}

@@

\noindent
Both \ZpLC and \ZpLF compute the absolute precision of \ttv (at
time $t$) as the smallest valuation of an entry of $C_{t,\ttv}$ 
defined as the column of $M_t$ corresponding to the variable \ttv.
This sounds revelant because, if we believe that the submodule $H_t
\subset E_t$ is supposed to encode the precision on the variables
in $\calV_t$, Proposition~\ref{prop:precision} applied with the 
projection onto the $\ttv$-component $\pi_\ttv : E_t \to \Qp e_\ttv$ 
indicates that a good candidate for the precision on $e_\ttv$ is 
$\pi_\ttv(H_t)$. Identifying $\Qp e_\ttv$ with $\Qp$, $\pi_\ttv(H_t)$ is 
the $\Zp$-module generated by the entries of $C_{t,\ttv}$; it is then 
equal to $p^N \Zp$ where $N$ is the smallest valuation of an entry of 
$C_{t,\ttv}$.

@@

Taking the compositum of all the functions $\tilde F$ (\emph{cf} 
Eq.~\eqref{eq:tildeF}) from time $0$ to $t$, we find that the execution 
of the session until time $t$ is modeled by a mathematical function
$\Phi_t : \calU_t \oplus \calK_t \to E_t$.
From the design of \ZpLC, we deduce further that there exists a 
vector $\underline k_t \in K_t$ such that:
$$\Phi_t(\underline u_t \oplus \underline k_t) = \underline x_t 
\quad \text{and} \quad
d\Phi_t (U_t \oplus K_t) = H_t$$
where the differential of $\Phi_t$ is taken at the point
$\underline u_t \oplus \underline k_t$.

@@

\noindent \textbf{Diffused digits of precision.}

>>>>

\newtheorem{theo}{Theorem}[section]
\newtheorem{lem}[theo]{Lemma}
\newtheorem{prop}[theo]{Proposition}
\newtheorem{cor}[theo]{Corollary}
\newtheorem{quest}[theo]{Question}
\newtheorem{conj}[theo]{Conjecture}
\theoremstyle{definition}
\newtheorem{rmk}[theo]{Remark}
\newtheorem{ex}[theo]{Example}
\newtheorem{deftn}[theo]{Definition}

@@

Both \ZpLC and \ZpLF compute the absolute precision of \ttv (at time 
$t$) as the smallest valuation of an entry of the column of $M_t$ 
corresponding to the variable \ttv. Alternatively, it can be defined as 
the unique integer $N$ for which $\pi_\ttv(H_t) = p^N \Zp$ where 
$\pi_\ttv : E_t \to \Qp$ takes a vector to its \ttv-coordinate.
This definition of the absolute precision sounds revelant because, if we 
believe that the submodule $H_t \subset E_t$ is supposed to encode the 
precision on the variables in $\calV_t$, 
Proposition~\ref{prop:precision} applied with the mapping $\pi_\ttv$ 
indicates that a good candidate for the precision on $e_\ttv$ is 
$\pi_\ttv(H_t) = p^N \Zp$.

@@

Taking the compositum of all the functions $\tilde F$ (\emph{cf} 
Eq.~\eqref{eq:tildeF}) from time $0$ to $t$, we find that the execution 
of the session until time $t$ is modeled by a mathematical function
$\Phi_t : \calU_t \oplus \calK_t \to E_t$.
From the design of \ZpLC, we deduce further that there exists a 
vector $\underline k_t \in K_t$ such that:
$$\Phi_t(\underline u_t \oplus \underline k_t) = \underline x_t 
\quad \text{and} \quad
d\Phi_t (U_t \oplus K_t) = H_t$$
where the differential of $\Phi_t$ is taken at the point
$\underline u_t \oplus \underline k_t$. 
Set $\Phi_{t,\ttv} = \pi_\ttv \circ \Phi_t$. It maps 
$\underline u_t \oplus \underline k_t$ to the \ttv-coordinate 
$x_{t,\ttv}$ of $\underline x_t$ and satisfies:
$$d\Phi_{t,\ttv} (U_t \oplus K_t) = \pi_\ttv(H_t) = p^N \Zp$$
where $N$ is the value returned by \verb?precision_absolute?. Thus, 
as soon as the assumptions of Proposition \ref{prop:precision} are
fulfilled, we derive
$\Phi_{t,\ttv} \big((\underline u_t + U_t) \oplus (\underline k_t + K_t)\big) = 
x_{t,\ttv} + p^N \Zp$.
Noting that $k_t \in K_t$, we finally get:
\begin{equation}
\label{eq:absprec}
\Phi_{t,\ttv} (\underline u_t + U_t) \subset
\Phi_{t,\ttv} \big((\underline u_t + U_t) \oplus K_t\big) = 
x_{t,\ttv} + p^N \Zp.
\end{equation}
The latter inclusion means that the computed value $x_{t,\ttv}$ is
accurate at precision $O(p^N)$, \emph{i.e.} that the output absolute
precision is correct.

Unfortunately, checking automatically the assumptions of Proposition 
\ref{prop:precision} in full generality seems to be difficult, though it 
can be done by hand for many particular examples \cite{}.

\begin{rmk}
Assuming that Proposition \ref{prop:precision} applies, 
the absolute precision computed as above is optimal if and only if
the inclusion of \eqref{eq:absprec} is an equality. Applying again
Proposition \ref{prop:precision}
with the restricted mapping $\Phi_{t,\ttv} : \calU_t \to \Qp$ and the
lattice $U_t$, we find that this happens if and only if
$d\Phi_{t,\ttv}(U_t) = p^N \Zp$.

Unfortunately, the latter condition cannot be checked on the matrix 
$M_t$ (because of reductions). 
However it is possible (and easy) to check whether the weaker condition 
$d\Phi_{t,\ttv}(K_t) \subsetneq p^N \Zp$. This checking is achieved by 
the method
\verb?is_precision_capped? (provided by our package) which returns true 
if $d\Phi_{t,\ttv}(K_t) = p^N \Zp$. As a consequence, when this method
answers false, the absolute precision computed by the software is
likely optimal.
\end{rmk}

@@

\noindent \textit{Diffused digits of precision.}
